---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 100%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE}
source('readData.r')
```


# By-POS Slopes for Surprisal and UID Measures
```{r, echo=FALSE}



# first extract the POS part of the dataset
POS_slopes = data %>% filter(!is.na(Var_Slope_Surp_Word)) 
POS_slopes = POS_slopes %>% select(ModelName,FileName, RegType, Language, POS, Mean_Slope_Surp_POS, Var_Slope_Surp_POS, Var_Slope_Surp_Word, Mean_Slope_Surp_Word, Var_Slope_UID, Mean_Slope_UID) %>% mutate(Surp_POS_t = Mean_Slope_Surp_POS / sqrt(abs(Var_Slope_Surp_POS)), Surp_Word_t = Mean_Slope_Surp_Word / sqrt(abs(Var_Slope_Surp_Word)), UID_t = Mean_Slope_UID / sqrt(abs(Var_Slope_UID)))

#POS_slopes[order(POS_slopes$Surp_POS_t),]

##############################
```

Error bars indicate estimated posterior standard deviations.
As the effects of all measures (surprisal, dependency length, etc.) are negative, positive values of the random slopes indicate a weakened effect for a given POS category.

## POS Surprisal
```{r, echo=FALSE, fig.width=10, fig.height=3}
for(language in unique(POS_slopes$Language)) {
dodge = position_dodge(.9)
plot = ggplot(POS_slopes %>% filter(Language==language), aes(x=POS,y=Mean_Slope_Surp_POS, fill=RegType, color=FileName)) +
  geom_errorbar(aes(ymin=Mean_Slope_Surp_POS-sqrt(abs(Var_Slope_Surp_POS)), ymax=Mean_Slope_Surp_POS+sqrt(abs(Var_Slope_Surp_POS))), position=dodge, width=.25) +
  geom_bar(stat="identity", position=dodge) +
  facet_wrap( ~ Language) 
print(plot)
}
```

## Word Surprisal
```{r, echo=FALSE, fig.width=10, fig.height=3}
dodge = position_dodge(.9)
for(language in unique(POS_slopes$Language)) {
plot = ggplot(POS_slopes %>% filter(Language==language), aes(x=POS,y=Mean_Slope_Surp_Word, fill=RegType, color=FileName)) +
  geom_errorbar(aes(ymin=Mean_Slope_Surp_Word-sqrt(abs(Var_Slope_Surp_Word)), ymax=Mean_Slope_Surp_Word+sqrt(abs(Var_Slope_Surp_Word))), position=dodge, width=.25) +
  geom_bar(stat="identity", position=dodge) +
  facet_wrap( ~ Language) 
print(plot)
}
```

## Squared Deviation of Suprisal
```{r, echo=FALSE, fig.width=10, fig.height=3}
dodge = position_dodge(.9)
for(language in unique(POS_slopes$Language)) {
plot = ggplot(POS_slopes %>% filter(Language==language), aes(x=POS,y=Mean_Slope_UID, fill=RegType, color=FileName)) +
  geom_errorbar(aes(ymin=Mean_Slope_UID-sqrt(abs(Var_Slope_UID)), ymax=Mean_Slope_UID+sqrt(abs(Var_Slope_UID))), position=dodge, width=.25) +
  geom_bar(stat="identity", position=dodge) +
  facet_wrap( ~ Language) 
print(plot)
}
```






