---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 150%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE, warning=FALSE, output=FALSE, include=FALSE}
source('readData.r')
```





```{r, echo=FALSE, warning=FALSE, output=FALSE, include=FALSE}
source('processRealWeights.r')
data = data %>% filter(Head != "PUNCT", Dependent != "PUNCT", CoarseDependency != 'punct')
data = data %>% filter(Counter > 20000)

```




```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossHeads = function(dataset) {
    for(language in unique(dataset$Language)) {
        dodge = position_dodge(.9)
        agr = dataset %>% filter(Language==language) %>% group_by(Language, Head) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE))
        plot = ggplot(agr, aes(x=Head,y=DistanceWeight, fill=Real_DistanceWeight)) +
          geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
          geom_bar(stat="identity", position=dodge) +
          facet_wrap(~Language)
        print(plot)
    }
}
```


```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossHeadsAveragedOverLangs = function(dataset) {
    dodge = position_dodge(.9)
    agr = dataset %>% group_by(Head) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE))
    plot = ggplot(agr, aes(x=Head,y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge)
    print(plot)
}
```


```{r, echo=FALSE, fig.width=13, fig.height=3}
generalOrderPredictionsAcrossHeads = function(dataset) {
  for(language in unique(dataset$Language)) {
      dodge = position_dodge(.9)
      agr = dataset %>% filter(Language==language) %>% group_by(Language,Head) %>% summarise(DH_Weight_SD = weighted.sd(DH_Weight, Count), DH_Weight = weighted.mean(DH_Weight, Count), Real_DH_Weight = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE))
      plot = ggplot(agr, aes(x=Head,y=DH_Weight, fill=Real_DH_Weight)) +
        geom_errorbar(aes(ymin=DH_Weight-DH_Weight_SD, ymax=DH_Weight+DH_Weight_SD), position=dodge, width=.25) +
        geom_bar(stat="identity", position=dodge) +
        facet_wrap(~Language)
      print(plot)
  }
}
```
```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependents = function(dataset) {
for(language in unique(dataset$Language)) {
    dodge = position_dodge(.9)
    agr = dataset %>% filter(Language==language) %>% group_by(Language, Dependent) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE))
    plot = ggplot(agr, aes(x=Dependent,y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) +
      facet_wrap(~Language)
    print(plot)
}
}
```



```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependentsAveragedOverLanguages = function(dataset) {
    dodge = position_dodge(.9)
    agr = dataset  %>% group_by(Dependent) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE))
    plot = ggplot(agr, aes(x=Dependent,y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge)
    print(plot)
}
```

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalOrderPredictionsAcrossDependents = function(dataset) {
  for(language in unique(dataset$Language)) {
      dodge = position_dodge(.9)
      agr = dataset %>% filter(Language==language) %>% group_by(Language,Dependent) %>% summarise(DH_Weight_SD = weighted.sd(DH_Weight, Count), DH_Weight = weighted.mean(DH_Weight, Count), Real_DH_Weight = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE))
      plot = ggplot(agr, aes(x=Dependent,y=DH_Weight, fill=Real_DH_Weight)) +
        geom_errorbar(aes(ymin=DH_Weight-DH_Weight_SD, ymax=DH_Weight+DH_Weight_SD), position=dodge, width=.25) +
        geom_bar(stat="identity", position=dodge) +
        facet_wrap(~Language)
      print(plot)
  }
}
```


```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependencies = function(dataset) {

for(language in unique(dataset$Language)) {
    dodge = position_dodge(.9)
    agr = dataset %>% filter(Language==language) %>% group_by(Language, Dependency) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE))
    plot = ggplot(agr, aes(x=Dependency,y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) +
      facet_wrap(~Language)
    print(plot)

    
    agr = dataset %>% filter(Language==language) %>% mutate(Direction = ifelse(DH_Weight > 0, "DH", "HD")) %>% group_by(Language, Direction, Dependency) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE))
    plot = ggplot(agr, aes(x=Dependency,y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) +
      facet_wrap(~Language+Direction)
    print(plot)
}
}
```

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependenciesAveraged = function(dataset) {
    dodge = position_dodge(.9)
    agr = dataset %>% group_by(Dependency) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), Count = sum(Count)) %>% filter(Count >= median(Count))
    plot = ggplot(agr, aes(x=Dependency,y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) 
    print(plot)
}
```


```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependenciesAveraged2 = function(dataset) {
    dodge = position_dodge(.9)
    agr = dataset %>% group_by(Head,Dependency,Dependent) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), Count = sum(Count))
    agr = agr[order(-agr$Count),] 
    agr = agr[(1:50),]
    plot = ggplot(agr, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge)
    print(plot)
}
```



```{r, echo=FALSE, fig.width=35, fig.height=3}
generalDistancePredictionsAcrossDependenciesAveraged3 = function(dataset) {
    dodge = position_dodge(.9)
    dataset = dataset %>% mutate(Direction = ifelse(DH_Weight > 0, "DH", "HD"))
    agr = dataset %>% group_by(Direction,Head,Dependency,Dependent) %>% summarise(DistanceWeight_SD = weighted.sd(DistanceWeight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Real_DistanceWeight = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), Count = sum(Count))
    agr = agr[order(-agr$Count),] 
    agr = agr[(1:50),]
    plot = ggplot(agr, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=DistanceWeight, fill=Real_DistanceWeight)) +
      geom_errorbar(aes(ymin=DistanceWeight-DistanceWeight_SD, ymax=DistanceWeight+DistanceWeight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) +
      facet_wrap( ~ Direction)
    print(plot)
}
```
























```{r, echo=FALSE, fig.width=13, fig.height=3}
generalOrderPredictionsAcrossDependencies = function(dataset) {
  for(language in unique(dataset$Language)) {
      dodge = position_dodge(.9)
      agr = dataset %>% filter(Language==language) %>% group_by(Language,Dependency) %>% summarise(DH_Weight_SD = weighted.sd(DH_Weight, Count), DH_Weight = weighted.mean(DH_Weight, Count), Real_DH_Weight = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE))
      plot = ggplot(agr, aes(x=Dependency,y=DH_Weight, fill=Real_DH_Weight)) +
        geom_errorbar(aes(ymin=DH_Weight-DH_Weight_SD, ymax=DH_Weight+DH_Weight_SD), position=dodge, width=.25) +
        geom_bar(stat="identity", position=dodge) +
        facet_wrap(~Language)
      print(plot)
  }
}
```



```{r, echo=FALSE, fig.width=35, fig.height=3}
generalOrderPredictionsAcrossDependenciesAveraged = function(dataset) {
    dodge = position_dodge(.9)
    agr = dataset %>% group_by(Dependency) %>% summarise(DH_Weight_SD = weighted.sd(DH_Weight, Count), DH_Weight = weighted.mean(DH_Weight, Count), Real_DH_Weight = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Count = sum(Count)) %>% filter(Count >= quantile(Count, 0.7))
    plot = ggplot(agr, aes(x=Dependency,y=DH_Weight, fill=Real_DH_Weight)) +
      geom_errorbar(aes(ymin=DH_Weight-DH_Weight_SD, ymax=DH_Weight+DH_Weight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge)
    print(plot)

    dodge = position_dodge(.9)
    agr = dataset %>% group_by(Head,Dependency,Dependent) %>% summarise(DH_Weight_SD = weighted.sd(DH_Weight, Count), DH_Weight = weighted.mean(DH_Weight, Count), Real_DH_Weight = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Count = sum(Count))
    agr = agr[order(-agr$Count),] 
    agr = agr[(1:50),]
    plot = ggplot(agr, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=DH_Weight, fill=Real_DH_Weight)) +
      geom_errorbar(aes(ymin=DH_Weight-DH_Weight_SD, ymax=DH_Weight+DH_Weight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge)
    print(plot)

    dodge = position_dodge(.9)
    dataset = dataset %>% mutate(Direction = ifelse(DH_Weight > 0, "DH", "HD"))
    agr = dataset %>% group_by(Direction,Head,Dependency,Dependent) %>% summarise(DH_Weight_SD = weighted.sd(DH_Weight, Count), DH_Weight = weighted.mean(DH_Weight, Count), Real_DH_Weight = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Count = sum(Count))
    agr = agr[order(-agr$Count),] 
    agr = agr[(1:50),]
    plot = ggplot(agr, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=DH_Weight, fill=Real_DH_Weight)) +
      geom_errorbar(aes(ymin=DH_Weight-DH_Weight_SD, ymax=DH_Weight+DH_Weight_SD), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) +
      facet_wrap( ~ Direction)
    print(plot)
}
```







# Minimizing Dependency Length

Here I'm plotting data from the counterfactual languages optimized to minimize dependency length.

'DistanceWeight' is the distance parameter in the ordering model. Higher values indicate greater distance between head and dependent. 


```{r, echo=FALSE}
dataDepL = data[data$Objective == "DepL",]
depLDistances = as.data.frame(dataDepL %>% group_by(Dependency) %>% summarise(DH_Weight = weighted.mean(DH_Weight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), Count=sum(Count)))

depLDistancesDetails = as.data.frame(dataDepL %>% group_by(Dependency, Language, FileName) %>% summarise(DH_Weight = weighted.mean(DH_Weight, Count, na.rm=TRUE), DistanceWeight = weighted.mean(DistanceWeight, Count, na.rm=TRUE), DH = weighted.mean(DH, Count, na.rm=TRUE), Count=sum(Count, na.rm=TRUE)))

depLPerHead = as.data.frame(dataDepL %>% group_by(Head,Dependency,  Language, FileName, AverageLoss) %>% summarise(DH_Weight = weighted.mean(DH_Weight, Count), DistanceWeight = weighted.mean(DistanceWeight, Count), DH = weighted.mean(DH, Count, na.rm=TRUE), Count=sum(Count)))

depLNP = depLPerHead %>% filter(Head == "NOUN") #%>% filter(Count > 1000)
depL_G_20 = depLNP %>% filter(Dependency %in% c("amod", "det", "nummod"))

# distance
#depL_G_20 %>% group_by(Dependency) %>% summarise(DistanceWeight = mean(DistanceWeight))

# look at existing patterns
#depL_G_20[order(depL_G_20$Language, depL_G_20$DH, depL_G_20$DistanceWeight),]

depL_G_20 = depL_G_20 %>% mutate(DH = DH_Weight>0)
```




## Distance logits for Det, Num, Adj
More negative values indicate stronger preference to be close to the noun.
Here, adjectives are predicted to occur at higher distance from the noun than determiners.

```{r, echo=FALSE}
plot = ggplot(depL_G_20, aes(DistanceWeight, fill=Language)) +
  geom_histogram(binwidth=2) +
  facet_wrap( ~ Dependency) 
print(plot)



library(data.table)
depL_G_20 = dcast(setDT(depL_G_20), Language +FileName ~ Dependency, value.var=c("DistanceWeight", "DH_Weight"))




##########

depLCoreDependents = depLPerHead %>% filter(Head == "VERB") %>% filter(Count > 1000)
depL_CoreDependents = depLCoreDependents %>% filter(Dependency %in% c("obj", "iobj", "nsubj", "nsubjpass"))

## distance
#depL_CoreDependents %>% group_by(Dependency) %>% summarise(DistanceWeight = mean(DistanceWeight))

## look at existing patterns
#depL_CoreDependents[order(depL_CoreDependents$Language, depL_CoreDependents$DH, depL_CoreDependents$DistanceWeight),]

```

### Distance logits for Core Arguments
More negative values indicate stronger preference to be close to the noun.
Subjects are predicted to come closer.

```{r, echo=FALSE}
plot = ggplot(depL_CoreDependents, aes(DistanceWeight, fill=Language)) +
  geom_histogram(binwidth=2) +
  facet_wrap( ~ Dependency) 
print(plot)



library(data.table)
depL_CoreDependents = dcast(setDT(depL_CoreDependents), Language +FileName ~ Dependency, value.var=c("DistanceWeight", "DH_Weight"), na.rm=TRUE)


#################################

## nominal vs pronominal core arguments

#lmPerHeadDep = as.data.frame(dataLM %>% group_by(Head,Dependency,  Language, FileName, AverageLoss) %>% summarise(DH_Weight = weighted.mean(DH_Weight, Count, na.rm=TRUE), DistanceWeight = weighted.mean(DistanceWeight, Count), DH = weighted.mean(DH, Count, na.rm=TRUE), Count=sum(Count)))


```

### Distance logits for Nominal and Pronominal Arguments
More negative values indicate stronger preference to be close to the noun.
Pronominal arguments are predicted to come closer than nominal arguments.
```{r, echo=FALSE}

depLNounPronArgs = dataDepL %>% filter(Head == "VERB", Dependent %in% c("NOUN", "PRON")) %>%group_by(Dependency, Dependent, Language, FileName, AverageLoss)  %>% summarise(DH_Weight = weighted.mean(DH_Weight, Count, na.rm=TRUE), DistanceWeight = weighted.mean(DistanceWeight, Count), DH = weighted.mean(DH, Count, na.rm=TRUE), Count=sum(Count))

depL_NounPronArgs = depLNounPronArgs %>% filter(Dependency %in% c("obj", "iobj", "nsubj", "nsubjpass"))

plot = ggplot(depL_NounPronArgs, aes(DistanceWeight, fill=Language)) +
  geom_histogram(binwidth=2) +
  facet_wrap( ~ Dependency + Dependent) 
print(plot)



library(data.table)
depL_NounPronArgs = dcast(setDT(depL_NounPronArgs), Language +FileName ~ Dependency + Dependent, value.var=c("DistanceWeight", "DH_Weight"), na.rm=TRUE)



## prepositions vs postpositions
prePostpos = dataDepL %>% filter(Head == "NOUN", Dependency == "case", Dependent == "ADP") %>%group_by(Dependency, Dependent, Language, FileName, AverageLoss)  %>% summarise(DH_Weight = weighted.mean(DH_Weight, Count, na.rm=TRUE), DistanceWeight = weighted.mean(DistanceWeight, Count), DH = weighted.mean(DH, Count, na.rm=TRUE), Count=sum(Count))


library(data.table)
prePostpos = dcast(setDT(prePostpos), Language +FileName ~ 1, value.var=c("DistanceWeight", "DH_Weight"), na.rm=TRUE)

prePostposByLan = prePostpos %>% group_by(Language) %>% summarise(DH_Weight_SD = sd(DH_Weight, na.rm=TRUE), DH_Weight = mean(DH_Weight, na.rm=TRUE), DistanceWeight = mean(DistanceWeight, na.rm=TRUE))
## cor(prePostposByLan$DH_Weight, prePostposByLan$DistanceWeight)
## `when before the head, bind more strongly'

#################################











#dataDepL = data[data$ModelName == "readDataDistEnglishGPUDepLength.py",]
#depLDistances = as.data.frame(dataDepL %>% group_by(Dependency) %>% summarise(DistanceWeight = weighted.mean(DistanceWeight, Count)))

#dataEnglish = data[data$ModelName == "inferWeightsForEnglish.py",]
#englishDistances = as.data.frame(dataEnglish %>% group_by(Dependency) %>% summarise(DistanceWeight = weighted.mean(DistanceWeight, Count)))


#data = merge(data, dataEnglish %>% rename(DistanceWeightEnglish=DistanceWeight,DH_WeightEnglish = DH_Weight) %>% select(Head, Dependency, Dependent, DistanceWeightEnglish, DH_WeightEnglish), by=c("Head", "Dependency", "Dependent"))

## [1] "Head"           "Dependency"     "Dependent"      "FileName"      
## [5] "ModelName"      "Quality"        "Direction"      "DistanceWeight"
## [9] "DH_Weight"      "Counter"        "Count"         


```



# Values for other Dependencies

Here I'm plotting dependent-first and distance logits obtained by optimizing for dependency length.

For the "Distance" logits, higher values mean greater distance from the head. 

For the "DH" (Dependent-First) logits, higher values mean greater preference for the dependent to come first.

As the number of dependency triples is large, I'm plotting averages along the dimensions head POS, dependency label, and dependent POS.
I start with by-dependent averages since they seem easiest to interpret.


## Averaged across Dependent POS
Here I'm plotting averages of inferred values by the POS of the dependent of each dependency triple.


### Averaged Distance Predictions across Dependents

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependents(dataDepL)
```

#### averaged over languages

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependentsAveragedOverLanguages(dataDepL)
```

### Averaged Order Predictions across Dependents

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalOrderPredictionsAcrossDependents(dataDepL)
```



### Averaged across Head POS
Here I'm plotting averages of inferred values, by the POS of the head of each dependency triple.


### Averaged Distance Predictions across Heads

There doesn't seem to be anything to see here.

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossHeads(dataDepL)
```

#### averaged over languages

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossHeadsAveragedOverLangs(dataDepL)
```

### Averaged Order Predictions across Heads

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalOrderPredictionsAcrossHeads(dataDepL)
```


## Averaged across Dependency Label
Here I'm plotting averages of inferred values by the label of the dependency.



### Averaged Distance Predictions across Dependencies

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependencies(dataDepL)
```

#### averaged across languages

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependenciesAveraged(dataDepL)
```

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependenciesAveraged2(dataDepL)
```

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalDistancePredictionsAcrossDependenciesAveraged3(dataDepL)
```

### Averaged Order Predictions across Dependencies

```{r, echo=FALSE, fig.width=13, fig.height=3}
generalOrderPredictionsAcrossDependencies(dataDepL)
```


#### averaged over Languages


```{r, echo=FALSE, fig.width=13, fig.height=3}
generalOrderPredictionsAcrossDependenciesAveraged(dataDepL)
```





