---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 250%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE}
source('readData.r')
```

```{r, echo=FALSE}
source('processRealWeights.r')
data = data %>% filter(Head != "PUNCT", Dependent != "PUNCT", CoarseDependency != 'punct')
data = data %>% filter(Counter > 20000)

```


# Random Slopes for Dependency Length


Throughout this section, colors indicate the degree to which the dependency is head-final in the language, measured as a logit (positive = more head-final than head-initial).

Finding: When controlling for surprisal (the `All' models), strong effects of Dependency Length occur mostly for head-initial dependencies.



```{r, echo=FALSE}
DepL_slopes = data %>% filter(!is.na(Mean_Slope_DepLength)) %>% select(PercentageOfLanguage, TotalCount, TotalCount_Dependent, Counter, RegType, ModelName, FileName, Language, FileName, Head, Dependency, CoarseDependency, Dependent, Mean_Slope_DepLength, Var_Slope_DepLength, Count, Real_DH_Mean, Real_Distance_Mean) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))
```



## Average Slopes by Dependent


```{r, echo=FALSE, fig.width=25, fig.height=5}
DepL_slopes_by_Dependent = DepL_slopes %>% group_by(Counter, RegType, ModelName, FileName, Language, FileName, Dependent) %>% summarise(Mean_Slope_DepLength = weighted.mean(Mean_Slope_DepLength, Count, na.rm=TRUE), Var_Slope_DepLength = weighted.mean(Var_Slope_DepLength, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE)) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))

DepL_slopes_by_Dependent = as.data.frame(DepL_slopes_by_Dependent)
#DepL_slopes_by_Dependent[order(DepL_slopes_by_Dependent$Slope_DepLength_t),]

#cor(DepL_slopes_by_Dependent$Real_DH_Mean, DepL_slopes_by_Dependent$Mean_Slope_DepLength)
#[1] 0.7085587
#DepL_slopesSig = DepL_slopes %>% filter(abs(Slope_DepLength_t) > 2)
#cor(DepL_slopesSig$Real_DH_Mean, DepL_slopesSig$Mean_Slope_DepLength)
#[1] 0.6987075
```

```{r, echo=FALSE, fig.width=15, fig.height=5}
#for(language in unique(DepL_slopes_by_Dependent$Language)) {
#  cat("\n###########\n")
#  cat(language)
#  cat("\n")
#  agg = DepL_slopes_by_Dependent %>% filter(Language==language)
#  agg = agg %>% filter(Counter > 20000)
# for(regtype in unique(agg$RegType)) {
#  agg2 = (agg %>% filter(RegType == regtype))
#  for(filename in unique(agg2$FileName)) {
#     agg3 = agg2 %>% filter(FileName == filename)
#     agg3 = agg3[order(agg3$Mean_Slope_DepLength),]
#     plot = ggplot(agg3, aes(x=Dependent,y=Mean_Slope_DepLength, fill=Real_DH_Mean)) +
#       geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
#       facet_wrap(~ Language + RegType + ModelName + FileName + Counter) +
#       geom_bar(stat="identity", position=dodge)
#     print(plot)
#     plot = ggplot(agg3, aes(x=Dependent,y=Mean_Slope_DepLength, fill=Real_Distance_Mean)) +
#       geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
#       facet_wrap(~ Language + RegType + ModelName + FileName + Counter) +
#       geom_bar(stat="identity", position=dodge)
#     print(plot)
#
#   }
#}
#}
##DepL_slopes[order(DepL_slopes$Slope_DepLength_t),]
##DepL_slopes = DepL_slopes %>% filter(!(FileName %in% c("4457600", "6037474")))
```


```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(DepL_slopes_by_Dependent$Language)) {
  agg = DepL_slopes_by_Dependent %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg3 = data.frame()
  for(filename in unique(agg$FileName)) {
        agg4 = agg %>% filter(FileName == filename)
        agg4 = agg4 %>% mutate(Var_Slope_DepLength = Var_Slope_DepLength / var(Mean_Slope_DepLength), Mean_Slope_DepLength = Mean_Slope_DepLength/sd(Mean_Slope_DepLength))
        agg3 = rbind(agg3, agg4)
  }
#  agg3 = agg3[order(-abs(agg3$Count)),]
#  agg3 = agg3[(1:min(nrow(agg3),15*length(unique(agg3$FileName)))),]
#  agg3 = agg3[order(-abs(agg3$Mean_Slope_DepLength)),]
#  agg3 = agg3[(1:min(nrow(agg3),30)),]
  plot = ggplot(agg3, aes(x=Dependent,y=Mean_Slope_DepLength, color=FileName, fill=Real_DH_Mean)) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)
  plot = ggplot(agg3, aes(x=Dependent,y=Mean_Slope_DepLength, color=FileName, fill=Real_Distance_Mean)) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)

  plot = ggplot(agg3, aes(x=Dependent,y=Mean_Slope_DepLength, color=FileName, fill=RegType)) +
    geom_bar(stat="identity", position=dodge) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) 
  print(plot)

}
#DepL_slopes[order(DepL_slopes$Slope_DepLength_t),]
#DepL_slopes = DepL_slopes %>% filter(!(FileName %in% c("4457600", "6037474")))
```

## Across Languages

```{r, echo=FALSE, fig.width=25, fig.height=5}

agg = rbind(DepL_slopes)

for(filename in unique(agg$FileName)) {
  belongs = (agg$FileName == filename)
  fileSD = sd(agg[belongs,]$Mean_Slope_DepLength)
  agg[belongs,]$Mean_Slope_DepLength = agg[belongs,]$Mean_Slope_DepLength/fileSD
}



agg = agg %>% group_by(Language, RegType, Dependent) %>% summarise(Mean_Slope_DepLength = weighted.mean(Mean_Slope_DepLength, Count, na.rm=TRUE), Var_Slope_DepLength = weighted.mean(Var_Slope_DepLength, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), TotalCount_Dependent = mean(TotalCount_Dependent)) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))

agg = as.data.frame(agg)

#bound = quantile(unique(agg$TotalCount_Dependent), 1 - 15.0/length(unique(agg$TotalCount_Dependent)), na.rm=TRUE)
#
#
#agg = agg %>% filter(TotalCount_Dependent >= bound)


  plot = ggplot(agg, aes(x=Dependent,y=Mean_Slope_DepLength, fill=Real_DH_Mean, group=paste(Language,RegType))) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)

  plot = ggplot(agg, aes(x=Dependent,y=Mean_Slope_DepLength, fill=Real_Distance_Mean, group=paste(Language,RegType))) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)

  plot = ggplot(agg, aes(x=Dependent,y=Mean_Slope_DepLength, fill=RegType, color=Language)) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)




```




