---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 250%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE}
source('readData.r')
```

```{r, echo=FALSE}
source('processRealWeights.r')
data = data %>% filter(Head != "PUNCT", Dependent != "PUNCT")
data = data %>% filter(Counter > 20000)

```


# Gradients for Dependency Length

Here I'm plotting approximate partial derivatives of expected total dependency length w.r.t. the parameters alpha und beta of the ordering model.



```{r, echo=FALSE}
WordSurp_grads = data %>% filter(ObjectiveName == "gradients_WordSurp") %>% select(PercentageOfLanguage, TotalCount, TotalCountCoarse, Counter, RegType, ModelName, FileName, Language, FileName, Head, Dependency, CoarseDependency, Dependent,   D_WordSurp_DH_Weight, D_WordSurp_DistanceWeight, Count, Real_DH_Mean, Real_Distance_Mean)
```



# Average Gradients by Dependency Triple

```{r, echo=FALSE}
WordSurp_grads_by_Dependency = WordSurp_grads %>% group_by(Counter, ModelName, FileName, Language, Head, Dependent, Dependency) # %>% summarise( = weighted.mean(Mean_Slope_WordSurpength, Count, na.rm=TRUE), Var_Slope_WordSurpength = weighted.mean(Var_Slope_WordSurpength, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), Count = sum(Count)) %>% mutate(Slope_WordSurpength_t = (Mean_Slope_WordSurpength / sqrt(Var_Slope_WordSurpength)))

WordSurp_grads_by_Dependency = as.data.frame(WordSurp_grads_by_Dependency)
```

## Gradients w.r.t. Dependent-First Logit

The 30 partial derivatives with the largest absolute value are shown.

More positive numbers mean that ordering this dependency as Dependent-First more frequently would increase average per-sentence aggregate surprisal.
For each language, the plot is colored by the actual Dependent-First Logit. 


```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(WordSurp_grads_by_Dependency$Language)) {
  agg = WordSurp_grads_by_Dependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg3 = agg[order(-abs(agg$D_WordSurp_DH_Weight)),]
  agg3 = agg3[(1:min(nrow(agg3),30)),]
  agg3 = agg3 %>% mutate(Triple = paste(Head, Dependency, Dependent, sep="\n"))
  plot = ggplot(agg3, aes(x=Triple,y=D_WordSurp_DH_Weight, color=FileName, fill=Real_DH_Mean)) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_WordSurp_DH_Weight, color=FileName, fill=Real_Distance_Mean)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_WordSurp_DH_Weight, color=FileName, fill=D_WordSurp_DistanceWeight)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
}
#WordSurp_grads[order(WordSurp_grads$Slope_WordSurpength_t),]
#WordSurp_grads = WordSurp_grads %>% filter(!(FileName %in% c("4457600", "6037474")))
```

## Gradients w.r.t. Distance Logit

The 30 partial derivatives with the largest absolute value are shown.

More positive numbers indicate that putting the dependent farther away from the head would increase average per-sentence surprisal.

```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(WordSurp_grads_by_Dependency$Language)) {
  agg = WordSurp_grads_by_Dependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg3 = agg[order(-abs(agg$D_WordSurp_DH_Weight)),]
  agg3 = agg3[(1:min(nrow(agg3),30)),]
  agg3 = agg3 %>% mutate(Triple = paste(Head, Dependency, Dependent, sep="\n"))
  plot = ggplot(agg3, aes(x=Triple,y=D_WordSurp_DistanceWeight, color=FileName, fill=Real_Distance_Mean)) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_WordSurp_DistanceWeight, color=FileName, fill=Real_DH_Mean)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_WordSurp_DistanceWeight, color=FileName, fill=D_WordSurp_DH_Weight)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
}
#WordSurp_grads[order(WordSurp_grads$Slope_WordSurpength_t),]
#WordSurp_grads = WordSurp_grads %>% filter(!(FileName %in% c("4457600", "6037474")))
```


# Findings

In some languages, there is a correlation between derivatives w.r.t. order and distance logits:
```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(WordSurp_grads_by_Dependency$Language)) {
  agg = WordSurp_grads_by_Dependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
print(cor.test(agg$D_WordSurp_DH_Weight, agg$D_WordSurp_DistanceWeight))
}
#WordSurp_grads[order(WordSurp_grads$Slope_WordSurpength_t),]
#WordSurp_grads = WordSurp_grads %>% filter(!(FileName %in% c("4457600", "6037474")))
```



On the other hand, there appear to be no correlations between the derivatives and the values of the logits.


