---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 100%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE}
source('readData.r')
```

```{r, echo=FALSE}
source('processRealWeights.r')
```

# Ordering Rules Controlling for Different Metrics

```{r, echo=FALSE}



dataResid = data[data$Objective == "residual",]
# cor(dataResid$DH_Mean, dataResid$Real_DH_Mean, use="complete")
#cor(dataResid$Distance_Mean, dataResid$Real_Distance_Mean, use="complete")

# sometimes Pyro sometimes gives negative variances,  well I should just transform them into something more stable myself
dataResid = dataResid %>% mutate(DH_Sigma = abs(DH_Sigma), Distance_Sigma = abs(Distance_Sigma), Real_DH_Sigma = abs(Real_DH_Sigma), Real_Distance_Sigma = abs(Real_Distance_Sigma))

dataResid = dataResid %>% mutate(DH_t = DH_Mean/sqrt(abs(DH_Sigma)), Distance_t = Distance_Mean/sqrt(abs(Distance_Sigma)), Real_DH_t = Real_DH_Mean/sqrt(abs(Real_DH_Sigma)), Real_Distance_t = Real_Distance_Mean/sqrt(abs(Real_Distance_Sigma)))
dataResid = dataResid %>% filter(!is.na(DH_Mean))
bigT = (dataResid[abs(dataResid$Distance_t - dataResid$Real_Distance_t) > 0.1,])

bigTFiltered = bigT[order(bigT$Distance_t),] %>% select(ModelName, RegType, Language, Head, Dependency, Dependent, Distance_t , Real_Distance_t)
#bigTFiltered = bigT[order(bigT$Distance_t),] %>% select(Head, Dependency, Dependent, Distance_Mean , Real_Distance_Mean)
```

## Distance Logits

### dependencies whose weights are `significant' according to only maximum-likelihood
Ordered by the quotient of t-values

```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(bigT$Language)) {
  agg = bigT %>% filter(Language==language, abs(Real_Distance_t) > 4)
  agg = agg[order(abs(agg$Distance_t/agg$Real_Distance_t)),]
  for(filename in unique(agg$FileName)) {
    agg2 = agg %>% filter(FileName == filename)
    agg2 = agg2[(1:min(30, nrow(agg2))),]

    agg2 = agg2 %>% mutate(Distance_Sigma = Distance_Sigma/var(Distance_Mean, na.rm=TRUE), Distance_Mean = Distance_Mean/sd(Distance_Mean, na.rm=TRUE))
    agg2 = agg2 %>% mutate(Real_Distance_Sigma = Real_Distance_Sigma/var(Real_Distance_Mean, na.rm=TRUE), Real_Distance_Mean = Real_Distance_Mean/sd(Real_Distance_Mean, na.rm=TRUE))

    agg3 = rbind(agg2) %>% mutate(Distance_Mean = Real_Distance_Mean, Distance_Sigma = Real_Distance_Sigma)
    agg3$RegType = "Baseline"
    agg2 = rbind(agg2, agg3)



    plot = ggplot(agg2, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=Distance_Mean, fill=RegType)) +
      geom_errorbar(aes(ymin=Distance_Mean-sqrt(abs(Distance_Sigma)),ymax=Distance_Mean+sqrt(abs(Distance_Sigma))), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) +
      facet_wrap( ~ Language + ModelName + FileName + Counter) 
    print(plot)
#    plot = ggplot(agg2, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=Real_Distance_Mean)) +
#      geom_errorbar(aes(ymin=Real_Distance_Mean-sqrt(abs(Real_Distance_Sigma)),ymax=Real_Distance_Mean+sqrt(abs(Real_Distance_Sigma))), position=dodge, width=.25) +
#      geom_bar(stat="identity", position=dodge) +
#      facet_wrap( ~ Language + ModelName + FileName + Counter) 
#    print(plot)
  }
}
```

### dependencies whose weights are `significant' according to only controlled model
```{r, echo=FALSE}
bigTFiltered %>% filter(abs(Distance_t) >= 4 & abs(Real_Distance_t) < 1)
```

### dependencies whose weights are more `significant' according to controlled model
Ordered by the quotient of t-values (opposite)

```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(bigT$Language)) {
  agg = bigT %>% filter(Language==language, abs(Real_Distance_t) > 4)
  agg = agg[order(-abs(agg$Distance_t/agg$Real_Distance_t)),]
  for(filename in unique(agg$FileName)) {
    agg2 = agg %>% filter(FileName == filename)
    agg2 = agg2[(1:min(nrow(agg2),30)),]

    agg2 = agg2 %>% mutate(Distance_Sigma = Distance_Sigma/var(Distance_Mean), Distance_Mean = Distance_Mean/sd(Distance_Mean))
    agg2 = agg2 %>% mutate(Real_Distance_Sigma = Real_Distance_Sigma/var(Real_Distance_Mean), Real_Distance_Mean = Real_Distance_Mean/sd(Real_Distance_Mean))


    agg3 = rbind(agg2) %>% mutate(Distance_Mean = Real_Distance_Mean, Distance_Sigma = Real_Distance_Sigma)
    agg3$RegType = "Baseline"
    agg2 = rbind(agg2, agg3)

    plot = ggplot(agg2, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=Distance_Mean, fill=RegType)) +
      geom_errorbar(aes(ymin=Distance_Mean-sqrt(abs(Distance_Sigma)),ymax=Distance_Mean+sqrt(abs(Distance_Sigma))), position=dodge, width=.25) +
      geom_bar(stat="identity", position=dodge) +
      facet_wrap( ~ Language + ModelName + FileName + Counter) 
    print(plot)
#    plot = ggplot(agg2, aes(x=paste(Head,Dependency,Dependent,sep="\n"),y=Real_Distance_Mean)) +
#      geom_errorbar(aes(ymin=Real_Distance_Mean-sqrt(abs(Real_Distance_Sigma)),ymax=Real_Distance_Mean+sqrt(abs(Real_Distance_Sigma))), position=dodge, width=.25) +
#      geom_bar(stat="identity", position=dodge) +
#      facet_wrap( ~ Language + ModelName + FileName + Counter) 
#    print(plot)
  }
}
```

## Direction Logits
```{r, echo=FALSE}


bigT = dataResid #(dataResid[abs(dataResid$DH_t - dataResid$Real_DH_t) > 1.0,])
#
#bigTFiltered = bigT[order(bigT$DH_t),] %>% select(RegType, Language, Head, Dependency, Dependent, DH_t , Real_DH_t)
##bigTFiltered = bigT[order(bigT$DH_t),] %>% select(Head, Dependency, Dependent, DH_Mean , Real_DH_Mean)
#
## show dependencies whose weights are `significant' according to only maximum-likelihood
#bigTFiltered %>% filter(abs(DH_t) < 1 & abs(Real_DH_t) >= 4)
#
## show dependencies whose weights are `significant' according to only the controlled model
#bigTFiltered %>% filter(abs(DH_t) >= 4 & abs(Real_DH_t) < 1)
#
## show dependencies whose weights are `significant' according to both
#bigTFiltered %>% filter(abs(DH_t) >= 4 & abs(Real_DH_t) >= 4)
#

# TODO control for the frequency of the different things. only want to say something about high-frequency dependencies.
##################

```


