---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 250%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE}
source('readData.r')
```

```{r, echo=FALSE}
source('processRealWeights.r')
data = data %>% filter(Head != "PUNCT", Dependent != "PUNCT")
data = data %>% filter(Counter > 20000)

```


# Random Slopes for Dependency Length


Throughout this section, colors indicate the degree to which the dependency is head-final in the language, measured as a logit (positive = more head-final than head-initial).

Finding: When controlling for surprisal (the `All' models), strong effects of Dependency Length occur mostly for head-initial dependencies.



```{r, echo=FALSE}
DepL_slopes = data %>% filter(!is.na(Mean_Slope_DepLength)) %>% select(TotalCount, TotalCountCoarse, Counter, RegType, ModelName, FileName, Language, FileName, Head, Dependency, CoarseDependency, Dependent, Mean_Slope_DepLength, Var_Slope_DepLength, Count, Real_DH_Mean, Real_Distance_Mean) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))
```

## Extreme Random Slopes
Strongly negative and strongly positive slopes (As before, negative =stronger effect of dependency length, positive=weaker effect):

```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(DepL_slopes$Language)) {
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg = DepL_slopes %>% filter(Language==language)
 for(regtype in unique(agg$RegType)) {
  cat(regtype)
  cat("\n")
  agg2 = (agg %>% filter(RegType == regtype))
  for(model in unique(agg2$ModelName)) {
     agg3 = agg2 %>% filter(ModelName == model)
     agg3 = agg3[order(agg3$Mean_Slope_DepLength),]
     agg3 = agg3 %>% mutate(X=paste(Head,Dependency,Dependent, sep="\n"))
     for(filename in unique(agg3$FileName)) {
       agg4 = agg3 %>% filter(FileName == filename)
       agg4$X = factor(agg4$X, levels=agg4$X)
       plot = ggplot(agg4[(1:min(nrow(agg4),30)),], aes(x=X,y=Mean_Slope_DepLength, fill=Real_DH_Mean)) +
         geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
         facet_wrap(~ Language + RegType + ModelName + FileName + Counter) +
         geom_bar(stat="identity", position=dodge)
       print(plot)

       plot = ggplot(agg4[(1:min(nrow(agg4),30)),], aes(x=X,y=Mean_Slope_DepLength, fill=Real_Distance_Mean)) +
         geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
         facet_wrap(~ Language + RegType + ModelName + FileName + Counter) +
         geom_bar(stat="identity", position=dodge)
       print(plot)
 
     
       plot = ggplot(agg4[(max(0,(nrow(agg4)-30)):nrow(agg4)),], aes(x=X,y=Mean_Slope_DepLength, fill=Real_DH_Mean)) +
         geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
         facet_wrap(~ Language + RegType + ModelName + FileName + Counter) +
         geom_bar(stat="identity", position=dodge)
       print(plot)
}
}
}
}
#DepL_slopes[order(DepL_slopes$Slope_DepLength_t),]
#DepL_slopes = DepL_slopes %>% filter(!(FileName %in% c("4457600", "6037474")))
```


