---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 150%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE, warning=FALSE, output=FALSE}
source('readData.r')
```

```{r, echo=FALSE, warning=FALSE, output=FALSE}
source('processRealWeights.r')
data = data %>% filter(Head != "PUNCT", Dependent != "PUNCT")
data = data %>% filter(Counter > 20000)

```


# Slopes for Dependency Length (by dependency triple)

Here, I show fitted per-dependency-triple slopes for dependency length, first by language and then for all languages together.
A more negative slope indicates above-average effect of dependency length.
Only the 15 most frequent triples are shown per language.
For each language, I show the plots colored by (1) the Dependent-First logit for the language, (2) the Distance logit, (3) the fixed effects that were included the model, among DepL (dependence length), surprisal, and ID (Variance of Surprisal (for UID)).

In general, slopes tend to become weaker when controlling for surprisal than when dependency length is the only fixed effect.

'FileName' indicates different runs of the model fitting procedure.


```{r, echo=FALSE}
DepL_slopes = data %>% filter(!is.na(Mean_Slope_DepLength)) %>% select(PercentageOfLanguage, TotalCount, TotalCountCoarse, Counter, RegType, ModelName, FileName, Language, FileName, Head, Dependency, CoarseDependency, Dependent, Mean_Slope_DepLength, Var_Slope_DepLength, Count, Real_DH_Mean, Real_Distance_Mean) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))
```



```{r, echo=FALSE}
DepL_slopes_by_Dependency = DepL_slopes %>% group_by(Counter, RegType, ModelName, FileName, Language, FileName, Head, Dependent, Dependency) %>% summarise(Mean_Slope_DepLength = weighted.mean(Mean_Slope_DepLength, Count, na.rm=TRUE), Var_Slope_DepLength = weighted.mean(Var_Slope_DepLength, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), Count = sum(Count)) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))

DepL_slopes_by_Dependency = as.data.frame(DepL_slopes_by_Dependency)
```



```{r, echo=FALSE, fig.width=15, fig.height=4}
for(language in unique(DepL_slopes_by_Dependency$Language)) {
  agg = DepL_slopes_by_Dependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg3 = data.frame()
  for(filename in unique(agg$FileName)) {
        agg4 = agg %>% filter(FileName == filename)
        agg4 = agg4 %>% mutate(Var_Slope_DepLength = Var_Slope_DepLength / var(Mean_Slope_DepLength), Mean_Slope_DepLength = Mean_Slope_DepLength/sd(Mean_Slope_DepLength))
        agg3 = rbind(agg3, agg4)
  }
#  threshold = quantile(unique(agg3$Count), 1-10/length(unique(agg3$Count)))
#  agg3 = filter(Count >= threshold)
  agg3 = agg3[order(-abs(agg3$Count)),]
  agg3 = agg3[(1:min(nrow(agg3),15*length(unique(agg3$FileName)))),]
  agg3 = agg3 %>% mutate(Triple = paste(Head, Dependency, Dependent, sep="\n"))
  plot = ggplot(agg3, aes(x=Triple,y=Mean_Slope_DepLength, color=FileName, fill=Real_DH_Mean)) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)
  plot = ggplot(agg3, aes(x=Triple,y=Mean_Slope_DepLength, color=FileName, fill=Real_Distance_Mean)) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)

  plot = ggplot(agg3, aes(x=Triple,y=Mean_Slope_DepLength, color=FileName, fill=RegType)) +
    geom_bar(stat="identity", position=dodge) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) 
  print(plot)

}
#DepL_slopes[order(DepL_slopes$Slope_DepLength_t),]
#DepL_slopes = DepL_slopes %>% filter(!(FileName %in% c("4457600", "6037474")))
```

# Across Languages



```{r, echo=FALSE, fig.width=15, fig.height=4}

agg = rbind(DepL_slopes)

for(filename in unique(agg$FileName)) {
  belongs = (agg$FileName == filename)
  fileSD = sd(agg[belongs,]$Mean_Slope_DepLength)
  agg[belongs,]$Mean_Slope_DepLength = agg[belongs,]$Mean_Slope_DepLength/fileSD
}

agg = agg %>% mutate(Triple = paste(Head, CoarseDependency, Dependent, sep="\n"))

# IMPOARTANT: cannot sum in this summarise, since this is collapsing across FileName, but also across fine-grained dependencies, so cannot necessarily replace sum by mean!
agg = agg %>% group_by(Language, RegType, Triple) %>% summarise(Mean_Slope_DepLength = weighted.mean(Mean_Slope_DepLength, Count, na.rm=TRUE), Var_Slope_DepLength = weighted.mean(Var_Slope_DepLength, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), TotalCountCoarse = mean(TotalCountCoarse)) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))

agg = as.data.frame(agg)

bound = quantile(unique(agg$TotalCountCoarse), 1 - 15.0/length(unique(agg$TotalCountCoarse)), na.rm=TRUE)


agg = agg %>% filter(TotalCountCoarse >= bound)


  plot = ggplot(agg, aes(x=Triple,y=Mean_Slope_DepLength, fill=Real_DH_Mean, group=paste(Language,RegType))) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)

  plot = ggplot(agg, aes(x=Triple,y=Mean_Slope_DepLength, fill=Real_Distance_Mean, group=paste(Language,RegType))) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)

  plot = ggplot(agg, aes(x=Triple,y=Mean_Slope_DepLength, fill=RegType, color=Language)) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)




```





