---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 250%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE}
source('readData.r')
```

```{r, echo=FALSE}
source('processRealWeights.r')
data = data %>% filter(Head != "PUNCT", Dependent != "PUNCT")
data = data %>% filter(Counter > 20000)

```


# Gradients for Dependency Length

Here I'm plotting approximate partial derivatives of expected total dependency length w.r.t. the parameters alpha und beta of the ordering model.



```{r, echo=FALSE}
DepL_grads = data %>% filter(ObjectiveName == "gradients_DepL") %>% select(PercentageOfLanguage, TotalCount, TotalCountCoarse, Counter, RegType, ModelName, FileName, Language, FileName, Head, Dependency, CoarseDependency, Dependent,   D_DepL_DH_Weight, D_DepL_DistanceWeight, Count, Real_DH_Mean, Real_Distance_Mean)
```



# Average Gradients by Dependency Triple

```{r, echo=FALSE}
DepL_grads_by_Dependency = DepL_grads %>% group_by(Counter, ModelName, FileName, Language, Head, Dependent, Dependency) # %>% summarise( = weighted.mean(Mean_Slope_DepLength, Count, na.rm=TRUE), Var_Slope_DepLength = weighted.mean(Var_Slope_DepLength, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), Count = sum(Count)) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))

DepL_grads_by_Dependency = as.data.frame(DepL_grads_by_Dependency)
```

## Gradients w.r.t. Dependent-First Logit

The 30 partial derivatives with the largest absolute value are shown.

More positive numbers mean that ordering this dependency as Dependent-First more frequently would increase average per-sentence dependency length.
For each language, the plot is colored by the actual Dependent-First Logit. 


```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(DepL_grads_by_Dependency$Language)) {
  agg = DepL_grads_by_Dependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg3 = agg[order(-abs(agg$D_DepL_DH_Weight)),]
  agg3 = agg3[(1:min(nrow(agg3),30)),]
  agg3 = agg3 %>% mutate(Triple = paste(Head, Dependency, Dependent, sep="\n"))
  plot = ggplot(agg3, aes(x=Triple,y=D_DepL_DH_Weight, color=FileName, fill=Real_DH_Mean)) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_DepL_DH_Weight, color=FileName, fill=Real_Distance_Mean)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_DepL_DH_Weight, color=FileName, fill=D_DepL_DistanceWeight)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
}
#DepL_grads[order(DepL_grads$Slope_DepLength_t),]
#DepL_grads = DepL_grads %>% filter(!(FileName %in% c("4457600", "6037474")))
```

## Gradients w.r.t. Distance Logit

The 30 partial derivatives with the largest absolute value are shown.

```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(DepL_grads_by_Dependency$Language)) {
  agg = DepL_grads_by_Dependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg3 = agg[order(-abs(agg$D_DepL_DH_Weight)),]
  agg3 = agg3[(1:min(nrow(agg3),30)),]
  agg3 = agg3 %>% mutate(Triple = paste(Head, Dependency, Dependent, sep="\n"))
  plot = ggplot(agg3, aes(x=Triple,y=D_DepL_DistanceWeight, color=FileName, fill=Real_Distance_Mean)) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_DepL_DistanceWeight, color=FileName, fill=Real_DH_Mean)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
#  plot = ggplot(agg3, aes(x=Triple,y=D_DepL_DistanceWeight, color=FileName, fill=D_DepL_DH_Weight)) +
#    facet_wrap(~ Language) +
#    geom_bar(stat="identity", position=dodge)
#  print(plot)
}
#DepL_grads[order(DepL_grads$Slope_DepLength_t),]
#DepL_grads = DepL_grads %>% filter(!(FileName %in% c("4457600", "6037474")))
```


# Findings

In some languages, there is a correlation between derivatives w.r.t. order and distance logits:
```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(DepL_grads_by_Dependency$Language)) {
  agg = DepL_grads_by_Dependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
print(cor.test(agg$D_DepL_DH_Weight, agg$D_DepL_DistanceWeight))
}
#DepL_grads[order(DepL_grads$Slope_DepLength_t),]
#DepL_grads = DepL_grads %>% filter(!(FileName %in% c("4457600", "6037474")))
```



On the other hand, there appear to be no correlations between the derivatives and the values of the logits.


