---
title: Exploring Word Order Predictions of Theories of Communicatively Efficient Language Production
author: Michael Hahn
date: March 2018
output:
  html_document:
    toc: true
---



<style>
img {
    max-width: 250%;

    /* other options:
    max-width: none;
    max-width: 700px;
    max-width: 9in;
    max-width: 25cm;
    etc
    */
}
</style>



```{r, echo=FALSE}
source('readData.r')
```

```{r, echo=FALSE}
source('processRealWeights.r')
data = data %>% filter(Head != "PUNCT", Dependent != "PUNCT", CoarseDependency != 'punct')
data = data %>% filter(Counter > 20000)

```


# Random Slopes for Dependency Length


Throughout this section, colors indicate the degree to which the dependency is head-final in the language, measured as a logit (positive = more head-final than head-initial).

Finding: When controlling for surprisal (the `All' models), strong effects of Dependency Length occur mostly for head-initial dependencies.



```{r, echo=FALSE}
DepL_slopes = data %>% filter(!is.na(Mean_Slope_DepLength)) %>% select(PercentageOfLanguage, TotalCount, TotalCount_CoarseDependency, Counter, RegType, ModelName, FileName, Language, FileName, Head, Dependency, CoarseDependency, Dependent, Mean_Slope_DepLength, Var_Slope_DepLength, Count, Real_DH_Mean, Real_Distance_Mean) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))
```



## Average Slopes by Dependency

Only 20 strongest coefficients shown.
```{r, echo=FALSE}
DepL_slopes_by_CoarseDependency = DepL_slopes %>% group_by(Counter, RegType, ModelName, FileName, Language, FileName, CoarseDependency) %>% summarise(Mean_Slope_DepLength = weighted.mean(Mean_Slope_DepLength, Count, na.rm=TRUE), Var_Slope_DepLength = weighted.mean(Var_Slope_DepLength, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE)) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))

DepL_slopes_by_CoarseDependency = as.data.frame(DepL_slopes_by_CoarseDependency)
```

```{r, echo=FALSE, fig.width=25, fig.height=5}
for(language in unique(DepL_slopes_by_CoarseDependency$Language)) {
  agg = DepL_slopes_by_CoarseDependency %>% filter(Language==language)
  cat("\n###########\n")
  cat(language)
  cat("\n")
  agg3 = data.frame()
  for(filename in unique(agg$FileName)) {
        agg4 = agg %>% filter(FileName == filename)
        agg4 = agg4 %>% mutate(Var_Slope_DepLength = Var_Slope_DepLength / var(Mean_Slope_DepLength), Mean_Slope_DepLength = Mean_Slope_DepLength/sd(Mean_Slope_DepLength))
        agg3 = rbind(agg3, agg4)
  }
  agg3 = agg3[order(-abs(agg3$Count)),]
  agg3 = agg3[(1:min(nrow(agg3),15*length(unique(agg3$FileName)))),]
#  agg3 = agg3[order(-abs(agg3$Mean_Slope_DepLength)),]
#  agg3 = agg3[(1:min(nrow(agg3),30)),]
  plot = ggplot(agg3, aes(x=CoarseDependency,y=Mean_Slope_DepLength, color=FileName, fill=Real_DH_Mean)) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)
  plot = ggplot(agg3, aes(x=CoarseDependency,y=Mean_Slope_DepLength, color=FileName, fill=Real_Distance_Mean)) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) +
    geom_bar(stat="identity", position=dodge)
  print(plot)

  plot = ggplot(agg3, aes(x=CoarseDependency,y=Mean_Slope_DepLength, color=FileName, fill=RegType)) +
    geom_bar(stat="identity", position=dodge) +
    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25) +
    facet_wrap(~ Language) 
  print(plot)

}
#DepL_slopes[order(DepL_slopes$Slope_DepLength_t),]
#DepL_slopes = DepL_slopes %>% filter(!(FileName %in% c("4457600", "6037474")))
```

## Across Languages

```{r, echo=FALSE, fig.width=25, fig.height=5}

agg = rbind(DepL_slopes)

for(filename in unique(agg$FileName)) {
  belongs = (agg$FileName == filename)
  fileSD = sd(agg[belongs,]$Mean_Slope_DepLength)
  agg[belongs,]$Mean_Slope_DepLength = agg[belongs,]$Mean_Slope_DepLength/fileSD
}



agg = agg %>% group_by(Language, RegType, CoarseDependency) %>% summarise(Mean_Slope_DepLength = weighted.mean(Mean_Slope_DepLength, Count, na.rm=TRUE), Var_Slope_DepLength = weighted.mean(Var_Slope_DepLength, Count, na.rm=TRUE), Real_DH_Mean = weighted.mean(Real_DH_Mean, Count, na.rm=TRUE), Real_Distance_Mean = weighted.mean(Real_Distance_Mean, Count, na.rm=TRUE), TotalCount_CoarseDependency = mean(TotalCount_CoarseDependency)) %>% mutate(Slope_DepLength_t = (Mean_Slope_DepLength / sqrt(Var_Slope_DepLength)))

agg = as.data.frame(agg)

bound = quantile(unique(agg$TotalCount_CoarseDependency), 1 - 15.0/length(unique(agg$TotalCount_CoarseDependency)), na.rm=TRUE)


agg = agg %>% filter(TotalCount_CoarseDependency >= bound)


  plot = ggplot(agg, aes(x=CoarseDependency,y=Mean_Slope_DepLength, fill=Real_DH_Mean, group=paste(Language,RegType))) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)

  plot = ggplot(agg, aes(x=CoarseDependency,y=Mean_Slope_DepLength, fill=Real_Distance_Mean, group=paste(Language,RegType))) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)

  plot = ggplot(agg, aes(x=CoarseDependency,y=Mean_Slope_DepLength, fill=RegType, color=Language)) +
    geom_bar(stat="identity", position=dodge) #+
#    geom_errorbar(aes(ymin=Mean_Slope_DepLength-sqrt(abs(Var_Slope_DepLength)),ymax=Mean_Slope_DepLength+sqrt(abs(Var_Slope_DepLength))), position=dodge, width=.25)
  print(plot)




```



